\begin{acknowledgements}

Sam Rigler, thank you, thank you, thank you for such good advice, delivered cold, direct and kind along with a recommendation that in the end may have changed everything.  I owe you.

Schoonover, math and models are joy.  You showed me that.  I didn't know it then.  I do now.

Pete, you're an ass.  But you got me into LLNL.  So, I suppose, thanks.  Ass.

Barry, you were my first (useful) mentor doing science not dancing around it in classes.  Thank you for the flurry of ideas mixed with long days of collection; It was what I dreamt science could be.

Daren, you took Sam's recommendation and then offered me the best possible job: a run of the place's instrumentation and a trust left alone I'd get shit done (and not break everything).  And when I wanted to leave anyway, you offered anything and everything you could do to help me out.  No words man, no words.

Carol, you took me on as student and I've still no idea why. You then let me loose to do things I had no certain sense how to do. I did them (wrong). I did them (less wrong). And now they're right. It wasn't a smooth path. The last five years though have been crucial, and you're how and why they happened. Thank you. Thank you. Thank you. Besides thank you, I'll say thank you by getting everything else out fast.

Max and Richard, your friendships are irreplacable.  Your confidence in me was misplaced, I think, for quite a while.  It was there though and I know it.  I'll fully deserve it one day, and the long work still needed to get there is how I'll pay my debt to you both.

Rachel, you make me do what I long to.  This is no small thing and I've no idea how you do it.  I am the kind-of-man-who-runs now.  With that everything else is attainable.  And you laugh with me, and at me.  You are so good, and right, and   grand.  I love you.

Mom, you've been unrelenting in support of my (sometimes changing) interests.  For that, endless thank yous.  I've never felt anything but love, for that I owe you too much to even start to make a list.  It's so much.  How about just this, you got me here.  This PhD is your fault.

Pop, you understood why I did this.  You really did. And you celebrated silently with me the risks I'll take now.  How I miss you.  Pop, this dissertation is yours.

\end{acknowledgements}